Self Regulation Ontology Retest Analyses

### 2018-03-20 code/SRO_Retest_Analyses.Rmd
refactoring

Instead of the summary metric like the retest reliability estimate for each measure we can check whether the difference score distribution for each measure

Make data frame with difference between two scores for each measure for each subject. Since the scores for different measures are on different scales for comparability the difference scores are scaled (i.e. divided by their root mean square) but not centered so a value of 0 for the difference scores would indicate a lack of a difference between the scores of a subject.

@@

But if I just run one mixed effects model then we don't get a sense of the simple effects (how many of the variables this effect of the days between is significant). I can run it separately for each dv to see if comparable proportions show a dependence on the delay for surveys and tasks.

```{r}
get_delay_effect = function(df){
  mod = lm(difference ~ scale(days_btw), data = df)
  out = data.frame(estimate=coef(summary(mod))["scale(days_btw)","Estimate"], pval=coef(summary(mod))["scale(days_btw)","Pr(>|t|)"])
  return(out)
}

@@

Instead I now plot all the data grouping by dependant variable and coloring depending on whether the difference score distribution for that variable has a significant slope when regressed over days between. The black is the main effect for the large multilevel model.

```{r}
t1_2_difference %>%
  mutate(sig_days_effect = ifelse(dv %in% sig_days_effect$dv, 1, 0))%>%
  arrange(-sig_days_effect) %>%
  ggplot()+
  stat_smooth (aes(as.numeric(days_btw), difference, 
                  group=factor(dv, levels=unique(dv[order(sig_days_effect)]), ordered=TRUE), 
                  color=factor(sig_days_effect, levels = c(1,0))),geom="line", alpha=0.5, method='lm')+
  geom_smooth(aes(as.numeric(days_btw), difference), method='lm', color = "black", se=FALSE)+
  theme_bw()+
  theme(legend.title = element_blank())+
  xlab('Days between completion')+
  ylab('Scaled difference score')+
  scale_color_discrete(breaks=c(0,1),
                       labels=c("NS", "Sig"))
```

```{r}
library(ggExtra)

tmp = t1_2_difference %>%
  filter(dv == "angling_risk_task_always_sunny.keep_adjusted_clicks")

p = ggplot(aes(as.numeric(days_btw), difference), data=tmp)+
  geom_point()+
  geom_smooth(method='lm')

ggMarginal(p, type='histogram')

p
```

```{r}
rm(sig_survey, sig_task, diff_from_zero_summary)
```

Significant fixed effect suggests that on average the longer the delay the larger the difference.

```{r}
summary(lmerTest::lmer(log(difference) ~ scale(days_btw)+(scale(days_btw) | dv), data=t1_2_difference %>% filter(dv != 'eating_survey'))) 
```

@@

>>>>

Instead of the summary metric like the retest reliability estimate for each measure we can check whether the difference score distribution for each measure depends on the delay between the two measurements. Since the difference score distribution is at subject level we can check whether the order of subjects in this distribution depends on their order in the distribution of days between completing the two tests.

Make data frame with difference between two scores for each measure for each subject. Since the scores for different measures are on different scales for comparability the difference scores are normalized (ie demeaned and dividied by the sd of the difference score distribution.) Note that the variance of the difference score distribution accounts for the variance in both time points by summing them. Normalization equates the means of each difference score distribution to 0 which would mask any meaningful change between the two time points but the analysis here does not interpret the mean of the difference score distributions but is interested in its relation to the days between completion. We check if the variables show systematic differences between the two points later and can investigate whether

Here we check if the difference is larger the longer the delay
What is the threshold for a meaningful difference between the two points for each variable? 
For standardized difference scores does it cross 2 on the y axis?

@@

How do the difference score distributions look like with respect to the days between completion?

```{r}
t1_2_difference %>%
  ggplot()+
  geom_smooth(aes(as.numeric(days_btw), abs(difference), group=factor(dv)), method='lm', se=FALSE)+
  geom_smooth(aes(as.numeric(days_btw), abs(difference)), method='lm', color = "black", se=FALSE)+
  theme_bw()+
  theme(legend.title = element_blank())+
  xlab('Days between completion')+
  ylab('Absolute Scaled difference score')

```

To test if the slope of the black is significant we would run a mixed effects model with a fixed effect for days between completion, random slope for each dv depending on the days between and random intercept for each dv.

Significant fixed effect suggests that on average the longer the delay the smaller the difference.

```{r}
summary(lmerTest::lmer(abs(difference) ~ scale(days_btw)+(scale(days_btw) | dv), data=t1_2_difference)) 
```

But if I just run one mixed effects model then we don't get a sense of the simple effects (how many of the variables this effect of the days between is significant and in which direction). I can run it separately for each dv to see if all difference score distributions are affected the same way depending on the days between completion.

```{r}
get_delay_effect = function(df){
  mod = lm(abs(difference) ~ scale(days_btw), data = df)
  out = data.frame(estimate=coef(summary(mod))["scale(days_btw)","Estimate"], pval=coef(summary(mod))["scale(days_btw)","Pr(>|t|)"])
  return(out)
}

@@

Instead I now plot all the data grouping by dependant variable and coloring depending on whether the difference score distribution for that variable has a significant slope when regressed over days between. The black is the main effect for the large multilevel model (this is the same plot as above colored by whether the difference score distribution for each dv has a significant slope depending on days between).

```{r}
t1_2_difference %>%
  mutate(sig_days_effect = ifelse(dv %in% sig_days_effect$dv, 1, 0))%>%
  arrange(-sig_days_effect) %>%
  ggplot()+
  stat_smooth (aes(as.numeric(days_btw), abs(difference), 
                  group=factor(dv, levels=unique(dv[order(sig_days_effect)]), ordered=TRUE), 
                  color=factor(sig_days_effect, levels = c(1,0))),geom="line", alpha=0.5, method='lm')+
  geom_smooth(aes(as.numeric(days_btw), abs(difference)), method='lm', color = "black", se=FALSE)+
  theme_bw()+
  theme(legend.title = element_blank())+
  xlab('Days between completion')+
  ylab('Scaled difference score')+
  scale_color_discrete(breaks=c(0,1),
                       labels=c("NS", "Sig"))
```

Conclusion: The average difference between the scores of a measure if anything decreases with the increased delay. This fixed effect masks the simple effect for each dv. When looked at individually only 10 variables show a dependence on days between completion. While 7 of them show decreasing differences depending on days between 3 do show increasing differences.
This suggests that the lack of stricter control over the days between completion of the measurement does not have a large effect on the stability of most measures.

effect of whether days between leads to a sig difference on reliability of that variable? plot coef of sig_day_effect over rel_df point estimate

@@

```{r}
sig_days_effect %>%
  left_join(rel_df, by='dv') %>%
  ggplot(aes(estimate, icc))+
  geom_point()+
  theme_bw()
```
